\newpage
\section*{Abstract}
The importance of the balance between exploration and exploitation plays a crucial role while solving the combinatorial optimization problems. This balance is reached by two general techniques: by using appropriate problem solver and by setting its proper parameters. Both problems were widely studied in the past and the research process continues up until now. The latest studies in the field of automated machine learning propose merging both problems, solving them at the design time and later strengthening the results in the runtime. To the best of our knowledge, the \emph{generalized} approach for solving the parameter setting in heuristic solvers has not yet been proposed. Therefore, the concept of merging heuristic selection and parameter control has not been introduced.

In this thesis we propose an approach for generic parameter control in meta-heuristics by means of reinforcement learning (RL). Making step further, we suggest a technique for merging the heuristic selection and parameter control problems and solving them in runtime using RL-based hyper-heuristic. The evaluation of proposed parameter control technique on symmetric traveling salesman problem (TSP) revealed its applicability by reaching the performance tuned in off-line underlying meta-heuristic. The concept of merging the heuristic selection and parameter control processes showed its ability to reach a performance of the best underling heuristic with tuned parameters on a subset of smaller TSP instances. However, it requires an implementation improvements for tackling a larger cases.